---
title: "Kaggle NY Taxi Duration"
output: html_notebook
---

## Libraries
```{r}
library(dplyr)
library(ggplot2)
library(caret)
library(MASS)
library(e1071)
library(neuralnet)
library(corrplot)
library(REdaS)
library(Metrics)
library(corrplot)
library(caTools)
```

## Data
```{r}
data = read.csv("train.csv")
```

```{r}
dataset = data
```

## Functions
### Getting distance
```{r}
haversine <- function(row_lon_lat) {
  
  pick_long = row_lon_lat[1]
  pick_lat = row_lon_lat[2]
  drop_long = row_lon_lat[3]
  drop_lat = row_lon_lat[4]
  
  R = 6371
  
  lon = apply(array(c(pick_long, drop_long)), MARGIN = 1, deg2rad)
  lat = apply(array(c(pick_lat, drop_lat)), MARGIN = 1, deg2rad)
  
  dlon = lon[2] - lon[1]
  dlat = lat[2] - lat[1]
  
  a = sin(dlat/2)^2 + cos(lat[1]) * cos(lat[2]) * sin(dlon/2)^2
  c = 2 * asin(sqrt(a))

  distance = R * c
}
```

### Remove and replace NA values UNUSED
```{r}
# data_rm.na = function(data){
#   value = 0
#   
#   mu = mean(data, na.rm = T)
#   sd = sd(data, na.rm = T)
#   
#   value = ifelse(is.na(data), as.integer(rnorm(1, mu, sd)), data)
#   value = ifelse(value < 0, value*-1, value)
# 
#   value
# }
```

### Calculate Pickup time
```{r}
get.time = function(data){
  
  len = length(data)
  temp = 0
  temp_date = 0
  temp_hour = 0
  time = 0
  
  # Splits date and time
  for (i in 1:len){
    temp[i] = strsplit(data[i], " ")
  }
  
  # Splits into smaller time units. 
  # i.e. hours, minutes, seconds
  for (i in 1:len){
    t = strsplit(temp[[i]][1], "-")
    t2 = strsplit(temp[[i]][2], ":")
    temp_date[i] = t[1]
    temp_hour[i] = t2[1]
  }
  
  # Converts time to seconds
  for (i in 1:len){
    time[i] = as.integer(temp_hour[[i]][1])*3600 + as.integer(temp_hour[[i]][2])*60 + as.integer(temp_hour[[i]][3])
  }
  
  # Returns array with time in seconds
  time
}
```

### Remove columns
```{r}
dataset = dataset %>%
  dplyr::select(-c(store_and_fwd_flag, id, vendor_id))
dataset
```

### Apply distance function to dataset
```{r}
# Temporal dataset to calculate distance
datatemp = dataset %>%
  dplyr::select(c(pickup_longitude, pickup_latitude, dropoff_longitude, dropoff_latitude))

# Get distance from latitude and longitude
# This takes time
datadistance = apply(
  datatemp,
  1, 
  haversine)

# Add distance to the original dataset
dataset$distance = datadistance

dataset 
```

### Calculate times from datetime values
```{r}
# Dropoff_time is not needed
# dropoff_time = get.time(dataset$dropff_datetime)
# dataset$dropoff_time = dropoff_time

# Get times in seconds
# This takes time
pickup_time = get.time(dataset$pickup_datetime)

# Add data to original dataset
dataset$pickup_time = pickup_time

# Only usefull data
dataset = dataset %>%
  dplyr::select(c(passenger_count, trip_duration, distance, pickup_time))
dataset
```
### Visualize trip_duration and pickup_time
```{r}
# This indicates skewed data graphically
ggplot(data = dataset, aes(trip_duration)) + geom_histogram()
```

### Skewness of the variables
```{r}
# skewness = 0  =>  perfect symemtry
# skewness <=  -1 || skewness >= 1  =>  highly skewed
# -1 <= skewness < -0.5 || 0.5 <= skewness < 1  =>  moderally skewed
# -0.5 <= skewness < 0 || 0 <= skewness < 0.5  =>  moderally symmetrical

skewness(dataset$trip_duration) # 343.1632 high af
skewness(dataset$pickup_time) # -0.44 just ok
skewness(dataset$distance) # 32.63091 high
skewness(dataset$passenger_count) # 2.1280 high
```


```{r}
# No aparent meaningful between both variables.
ggplot(dataset, aes(pickup_time, trip_duration)) + geom_point()
```

````{r}
ggplot(data = dataset, aes(x = pickup_time, y = trip_duration)) +
  geom_point() +
  scale_x_log10() + scale_y_log10()
```

### Min-Max Normalization UNUSED
```{r}
# temp_trip = dataset$trip_duration
# temp_pick = dataset$pickup_time
# 
# max = max(dataset$pickup_time)
# min = min(dataset$pickup_time)
# temp = 0
# for (i in 1:length(dataset$pickup_time)){
#   temp[i] = (dataset$pickup_time[i] - min) / (max - min)
# }
# dataset$pickup_time = temp
# 
# dataset
```


## Training 
### Remove NA values
```{r}
# Omit all NA data
dataset = na.omit(dataset)
```

### Data partition
```{r}
split = sample.split(dataset, SplitRatio = 0.7)
train_dataset = subset(dataset, split == "TRUE")
test_dataset = subset(dataset, split == "FALSE")
```

### Linear Models
```{r}
# Variable to predict
trip_duration = train_dataset$trip_duration

# Variables for prediction
distance = train_dataset$distance
passenger_count = train_dataset$passenger_count
pickup_time = train_dataset$pickup_time
```

### Normal linear model
```{r}
linear_model = train_dataset %>% lm(formula = trip_duration ~ distance + passenger_count + pickup_time)
```

### Log Transform linear models
```{r}
# log_linear_model1 = train_dataset %>% lm(formula = log1p(trip_duration) ~ distance + passenger_count + pickup_time)
# log_linear_model3 = train_dataset %>% lm(formula = log1p(trip_duration) ~ log1p(distance) + log1p(passenger_count) + pickup_time)
# log_linear_model4 = train_dataset %>% lm(formula = trip_duration ~ log1p(distance) + passenger_count + pickup_time)

# Best Log model based on the summaries
log_linear_model = train_dataset %>% lm(formula = log1p(trip_duration) ~ log1p(distance) + passenger_count + pickup_time)
```


### Summaries
```{r}
# summ1 = summary(log_linear_model1)
# summ3 = summary(log_linear_model3)
# summ4 = summary(log_linear_model4)

summ = summary(linear_model)
summ2 = summary(log_linear_model)

# summ1
# summ2
# summ3
summ 
summ2
```

### Make predictions
```{r}
duration_pred = predict(object = linear_model, newdata = test_dataset)
error = rmsle(test_dataset$trip_duration, duration_pred)
train_dataset$trip_duration
# This should be at least 0.4 or lower
error
```

## Test
### Apply best linear model to test data
```{r}
dataset_test = read.csv("test.csv")
dataset_test

id = dataset_test$id

dataset_test = dataset_test %>%
  dplyr::select(-c(store_and_fwd_flag, id, vendor_id))
dataset_test
```

### Get distance
```{r}
datatemp = dataset_test %>%
  dplyr::select(c(pickup_longitude, pickup_latitude, dropoff_longitude, dropoff_latitude))
# datatemp

# Esto toma tiempo 
datadistance = apply(
  datatemp,
  1, 
  haversine)

dataset_test$distance = datadistance

dataset_test = dataset_test %>%
  dplyr::select(-c(pickup_longitude, pickup_latitude, dropoff_longitude, dropoff_latitude))
dataset_test
```

### Pickup time-hour
```{r}
pickup_time = 0
pickup_time = get.time(dataset_test$pickup_datetime)
dataset_test$pickup_time = pickup_time

dataset_test = dataset_test %>%
  dplyr::select(-c(pickup_datetime))

dataset_test
```

### Make predictions
```{r}
duration_pred = predict(object = linear_model, newdata = dataset_test)
log_duration_pred = predict(object = log_linear_model, newdata = dataset_test)

head(duration_pred)
head(log_duration_pred)

# log_duration_pred = apply(array(log_duration_pred), 1, exp)

duration_pred = apply(array(duration_pred), 1, round)
log_duration_pred = apply(array(log_duration_pred), 1, round)

head(duration_pred)
head(log_duration_pred)
```


### Construct out-dataframe
```{r}
out_dataframe = data.frame(id = id, trip_duration = duration_pred)
out_dataframe
```

### Save data in Kaggle format for the linear model
```{r}
write.csv(out_dataframe, "kaggle_data.csv", quote = F, row.names = F)
```

## RMSLE
Lowest: 0.68607




























